\chapter{Introduction}\label{ch:introduction}

\section{Motivation}\label{sec:motivation}

Computer hardware and software continuously grow in size and complexity and as a
result ensuring the absence of exploitable behaviors is becoming increasingly
difficult. In the era in which computer systems are used extensively to
carry important information (e.g. credit card numbers, national security
documents), it has been widely accepted that security of these systems is a
priority. Researchers have identified a number of potential vulnerabilities
which arise from the violation of known but in-practice unenforceable safety
and security policies.

So far, computer security has been delegated mostly to software, while
the hardware is being almost completely controlled by the software.
High-level languages are becoming more widely used, due to features
such as strong type systems with type inference and automatic memory
management, making programming less error prone and reducing the
number of exploitable bugs. Furthermore, in order to strengthen the
security of computing systems a variety of low-level mitigation
techniques \cite{Cowan:1998:SAA:1267549.1267554,PaX,Erlingsson07}
have been proposed, however these are mostly ad-hoc solutions designed
to prevent specific known attacks, rather than enforcing a security
policy by preventing a well-defined class of attacks, thus making it
hard to reason about their effectiveness. In fact most of these
mitigation techniques can be circumvented by attackers
\cite{Szekeres2013}, which has lead to a continuous ``chase'' between
attackers and security researchers.

One common attack technique is to exploit some low-level vulnerability
such as a buffer overflow, in order to redirect the control flow to
attacker injected code. This attack can be stopped by a simple
protection scheme known as $W \oplus X$, which enforces that a memory
page is either executable or writable but not both.  Unfortunately,
clever attack techniques can bypass $W \oplus X$. In particular,
attackers have been using code-reuse attacks (e.g. return/jump -
oriented programming) that allows them to chain together existing
pieces of code to achieve malicious behavior without directly
introducing new code.  Abadi~\ETAL\cite{abadi2005} introduced a
property called Control Flow Integrity (CFI), which provides effective
protection against control-flow hijacking attacks. CFI enforces that
any execution of a program will respect a statically computed control
flow graph (CFG), thus thwarting all attacks that attempt to alter
the control-flow of a program, irregardless if the attacker
tries to redirect the control-flow to attacker injected code or
to an existing piece of code. \ch{missing references throughout}

\section{Contributions}\label{sec:contributions}
%could also be just a \paragraph

The main contribution of this thesis is the formalization and
verification in Coq of a dynamic monitor enforcing CFI, based on a
generic hardware-software security mechanism.

To this end, we used Coq to model a powerful attacker (\IE able to
execute buffer overflows) and proved that the mechanism enforces
\CFI even in the presence of such an attacker. In particular
we proved a variant of the \CFI property proposed by Abadi~\ETAL
\cite{AbadiBEL09}.

We managed to avoid tackling a direct and complex proof of this
theorem, by first defining an abstract machine that has \CFI by
construction, proving a simulation between the concrete and the
abstract machine and then transfering the \CFI property from the
abstract to the concrete level through a generic preservation theorem
that states that under certain assumptions \CFI is preserved by
backward simulation.

Additionally, we proved a two-way refinement, between a concrete
machine running the \CFI dynamic monitor and the abstract machine that
has \CFI by construction and acts as a specification to \CFI, showing
that the concrete machine emulates all behaviors of the
``correct-by-design'' abstract machine.


\section{Thesis Outline}\label{sec:outline}
\Cref{ch:policies} of this thesis briefly describes the motivation for
effective and efficient security policies, the desired properties a
robust security policy must satisfy and puts into context the
framework we utilize in order to formalize the Control-Flow Integrity policy
and reason about the effectiveness of the enforcement mechanism we used.

\Cref{ch:cfi} discusses the current state of
research on enforcing and formalizing Control-Flow Integrity and clarifies
the design choices of our approach regarding enforcement of \CFI.

\Cref{ch:verified_cfi} explains how we used the framework of \cref{ch:policies}
in order to formally reason about the security properties of the \CFI policy and
our approach to enforcing it.

Chapter 5.. conclusions, future work? Appendix with code and/or step
relations etc.?

\section{What needs to be done}

\begin{enumerate}
\item Take care of first parts (abstract, thanksgiving, keywords, outline, etc.)
  \ch{the abstract is the most important part of your thesis (\IE
    the part everyone will read), so start working on it soon,
    so that you can iterate a couple of times.}
\item \ch{The intro is a good start, but still quite weak: it gives
    only some motivation, but it's explaining only very briefly what
    this thesis is about (more like an abstract than an intro). One
    way to improve this is to work on your talk, and bring that
    high-level explanation into the intro (a couple more pages, with
    diagrams etc).}
\item Re-read and polish the whole thing
\item Optimize figure placement - once comments are removed and content is settled
  \ch{I almost never did this; try not to over-constrain LaTeX and it will
    usually do a good job; for large figures setting p allows LaTeX to
    display them on a whole page, which is usually a good idea. I've done
    this for Fig 2.2. and it worked great.}
  \nick{I agree especially since this is going to be printed and turned into a ``book''
    but what's really bothering latex is one line of text and then figure/listing
    and this happens a lot in chapter 4}
\item mention types on tags and DATA tag on registers
\item More things on concrete preservation?
\item A summary on the conclusions?
\item Call-stack protection in future work
\item Have a look at latest related work
\item think about appendix if we need one
\item think about diagrams, do we want more (\EG stopping for concrete machine)
\item Unified numbering? Theorems,figures,table all having one counter. Last time
  I tried to do this it failed. And it seems strange!
  \ch{I think theorems can have different counter from figures.
    I would call tables also figures; or if you meat listings, those
    can have a separate counter.
  I would use the same counter for
  theorems, lemmas, definitions, etc.}
\item Take care of first parts (abstract, thanksgiving, keywords, outline, etc.)
\item Decide capitalization. Symbolic or symbolic machine, enforce it through out the document.
\end{enumerate}